### 2023-08-20 Manuscript/sup.tex
sup

\documentclass[aip,jcp,preprint,noshowkeys,superscriptaddress]{revtex4-1}
\usepackage{graphicx,dcolumn,bm,xcolor,microtype,multirow,amscd,amsmath,amssymb,amsfonts,physics,wrapfig,txfonts,setspace}
\usepackage{siunitx}[=v2]
\usepackage[version=4]{mhchem}
%\usepackage{natbib}
%\bibliographystyle{achemso}

\maxdeadcycles=200

\newcommand{\ie}{\textit{i.e.}}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\alert}[1]{\textcolor{black}{#1}}
\usepackage[normalem]{ulem}
\newcommand{\titou}[1]{\textcolor{red}{#1}}
\newcommand{\trashPFL}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\PFL}[1]{\titou{(\underline{\bf PFL}: #1)}}
\newcommand{\toto}[1]{\textcolor{green}{#1}}
\newcommand{\trashAS}[1]{\textcolor{green}{\sout{#1}}}
\newcommand{\AS}[1]{\toto{(\underline{\bf AS}: #1)}}
\newcommand{\ant}[1]{\textcolor{orange}{#1}}
\newcommand{\SupInf}{\textcolor{blue}{Supporting Information}}

\newcommand{\mc}{\multicolumn}
\newcommand{\fnm}{\footnotemark}
\newcommand{\fnt}{\footnotetext}
\newcommand{\tabc}[1]{\multicolumn{1}{c}{#1}}
\newcommand{\QP}{\textsc{quantum package}}

\newcommand{\Ndet}{N_\text{det}}

\newcommand{\EHF}{E_\text{HF}}
\newcommand{\EDOCI}{E_\text{DOCI}}
\newcommand{\EFCI}{E_\text{FCI}}

\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\theequation}{S\arabic{equation}}

\usepackage[
	colorlinks=true,
    citecolor=blue,
    breaklinks=true
	]{hyperref}
\urlstyle{same}

\begin{document}

\newcommand{\LCPQ}{Laboratoire de Chimie et Physique Quantiques (UMR 5626), Universit\'e de Toulouse, CNRS, UPS, France}

\title{Supporting Information for ``Seniority and hierarchy configuration interaction for radicals and excited states''}

\author{F\'abris Kossoski}
\email{fkossoski@irsamc.ups-tlse.fr}
\affiliation{\LCPQ}
\author{Pierre-Fran\c{c}ois Loos}
\email{loos@irsamc.ups-tlse.fr}
\affiliation{\LCPQ}

% Abstract
\begin{abstract}
%Here comes the abstract.
%\bigskip
%\begin{center}
%        \boxed{\includegraphics[width=0.4\linewidth]{TOC}}
%\end{center}
%\bigskip
\end{abstract}

% Title
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Equilibrium geometry of ethylene}
%\label{sec:ethylene}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Equilibrium geometry of ethylene, in atomic units, obtained from Ref. \onlinecite{Loos_2018}:
\begin{singlespace}
\begin{verbatim}
C    0.00000000       1.26026583        0.00000000 
C    0.00000000      -1.26026583        0.00000000 
H    0.00000000       2.32345976        1.74287672 
H    0.00000000      -2.32345976        1.74287672 
H    0.00000000       2.32345976       -1.74287672 
H    0.00000000      -2.32345976       -1.74287672 
\end{verbatim}
\end{singlespace}

Equilibrium geometry of vinyl, in atomic units, obtained from Ref. \onlinecite{Loos_2020}:
\begin{singlespace}
\begin{verbatim}
C    0.00000000       1.16769663       -0.04303146
C    0.00000000      -1.29945364        0.15810072
H    0.00000000       2.38429609        1.59801822
H    0.00000000       2.08759130       -1.87998309
H    0.00000000      -2.90307925       -1.08814513
\end{verbatim}
\end{singlespace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{\ce{Computational details}}
%\label{sec:comp_details}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

By fitting the computed potential energy curves with a Morse potential, we obtained harmonic vibrational frequencies and equilibrium geometries.
The following intervals have been considered for the fitting:
\begin{itemize}
\item \ce{HF}: \SI{0.8}{\angstrom} to \SI{1.3}{\angstrom}
\item \ce{F2}: \SI{1.25}{\angstrom} to \SI{1.65}{\angstrom}
\item ethylene: \SI{2.2}{\bohr} to \SI{2.9}{\bohr}
\item \ce{N2}: \SI{0.95}{\angstrom} to \SI{1.3}{\angstrom}
\item \ce{H4}: \SI{1.45}{\bohr} to \SI{1.95}{\bohr}
\item \ce{H8}: \SI{1.6}{\bohr} to \SI{2.05}{\bohr}
\item \ce{OH}: \SI{0.85}{\angstrom} to \SI{1.2}{\angstrom}
\item \ce{CN}: \SI{1.0}{\angstrom} to \SI{1.4}{\angstrom}
\item vinyl: \SI{2.35}{\bohr} to \SI{2.7}{\bohr}
\item \ce{H7}: \SI{1.6}{\bohr} to \SI{2.05}{\bohr}
\end{itemize}

The non-parallelity error and distance error were computed for the following intervals:
\begin{itemize}
\item \ce{HF}: \SI{0.5}{\angstrom} to \SI{6}{\angstrom}
\item \ce{F2}: \SI{0.95}{\angstrom} to \SI{8}{\angstrom}
\item ethylene: \SI{1.5}{\bohr} to \SI{16}{\bohr}
\item \ce{N2}: \SI{0.7}{\angstrom} to \SI{4}{\angstrom}
\item \ce{H4}: \SI{1}{\bohr} to \SI{10}{\bohr}
\item \ce{H8}: \SI{1}{\bohr} to \SI{10}{\bohr}
\item \ce{OH}: \SI{0.6}{\angstrom} to \SI{8}{\angstrom}
\item \ce{CN}: \SI{0.7}{\angstrom} to \SI{4}{\angstrom}
\item vinyl: \SI{1.7}{\bohr} to \SI{9}{\bohr}
\item \ce{H7}: \SI{1}{\bohr} to \SI{10}{\bohr}
\end{itemize}

\clearpage

%%% FIG S1 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pes}
\caption{
Potential energy curves for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
according to FCI (black), hCI (green) (dashed lines for half-integer $h$), eCI (red), and sCI (blue) models, with the cc-pVDZ basis set and frozen-core approximation.}
\label{fig:plot_pes}
\end{figure*}
%%% %%% %%%

%%% FIG S2 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_OH}
\caption{
Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{OH},
as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
The dashed lines represent the FCI results.}
\label{fig:plot_pt2_rpt2_oh}
\end{figure*}
%%% %%% %%%

%%% FIG S3 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_CN}
\caption{
Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{CN},
as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
The dashed lines represent the FCI results.}
\label{fig:plot_pt2_rpt2_cn}
\end{figure*}
%%% %%% %%%

%%% FIG S4 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_vinyl}
\caption{
Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for vinyl,
as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
The dashed lines represent the FCI results.}
\label{fig:plot_pt2_rpt2_vinyl}
\end{figure*}
%%% %%% %%%

%%% FIG S5 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_H7}
\caption{
\label{fig:plot_pt2_rpt2_h7}
Non-parallelity error (NPE), distance error, equilibrium distance, and force constant, for \ce{H7},
as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
The dashed lines represent the FCI results.}
\end{figure*}
%%% %%% %%%

%%% FIG S6 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_HF}
\caption{
Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{HF},
as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
The dashed lines represent the FCI results.}
\label{fig:plot_pt2_rpt2_hf}
\end{figure*}
%%% %%% %%%

%%% FIG S7 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_F2}
\caption{
Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{F2},
as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
The dashed lines represent the FCI results.}
\label{fig:plot_pt2_rpt2_f2}
\end{figure*}
%%% %%% %%%

%%% FIG S8 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_ethylene}
\caption{
Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for ethylene,
as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
The dashed lines represent the FCI results.}
\label{fig:plot_pt2_rpt2_ethylene}
\end{figure*}
%%% %%% %%%

%%% FIG S9 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_N2}
\caption{
Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{N2},
as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
The dashed lines represent the FCI results.}
\label{fig:plot_pt2_rpt2_n2}
\end{figure*}
%%% %%% %%%

%%% FIG S10 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_H4}
\caption{
Non-parallelity error (NPE), distance error, equilibrium distance, and force constant, for \ce{H4},
as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
The dashed lines represent the FCI results.}
\label{fig:plot_pt2_rpt2_h4}
\end{figure*}
%%% %%% %%%

%%% FIG S11 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_H8}
\caption{
Non-parallelity error (NPE), distance error, equilibrium distance, and force constant, for \ce{H8},
as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
The dashed lines represent the FCI results.}
\label{fig:plot_pt2_rpt2_h8}
\end{figure*}
%%% %%% %%%

\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{manuscript}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}